---
title: "Data Task"
author: "Yiyu Chen"
date: "Jan 8th, 2021"
output: word_document
geometry: left = 1.5cm, right = 1.5cm, top = 1.5cm, bottom = 1.5cm
---
<style>
  body {font-size: 11pt;}
</style>
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
###load library
library(readxl)
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(data.table)
```


```{r, laod and clean data, echo=FALSE, include = F, message= F, warning=F}

###load data
mydata = read_excel(path = "./Test_Data.xlsx", sheet = "Data")

###Task 0: create csv with only one heading line
mydata = mydata[-1:-2,]
write.csv(mydata, file = "Data with one heading line.csv") 

###simple data cleaning
mydata= subset(mydata, !is.na(feelings_1))#exclude data with NA in feeling scores
#names(which(sapply(mydata, anyNA))) #check if there is NA data in feeling scores
mydata$ResponseId = as.factor(mydata$ResponseId) #change id to a factor
mydata[, 30:35] = lapply(mydata[,30:35], as.numeric)#change feelings to numeric variables
```



#### Task 1: create a single bar graph compare means of different feelings   

```{r, echo=FALSE}
###change to a long format
ix = c(9, 30:35, 67) #select relevant columns in the original data
temp = mydata[,ix] #create a temporary data frame with relevant columns
mydata.long = gather(temp, feeling, score, feelings_1:feelings_6, factor_key=TRUE)#change to a long format

###calculate relevant stat
plot.data = mydata.long %>%
  group_by(feeling) %>%
  summarise(mean = mean(score, na.rm=TRUE),
  sd = sd(score, na.rm=TRUE),
  n = n(),
  se=sd/sqrt(n),
  ci = qt(0.975, df=n-1)*se)

###visualization
ggplot(plot.data, aes(x=feeling, y=mean, group = factor(1))) + 
  geom_bar(stat="identity") +
  #error bar
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.1) +
  #text of titles
  labs(x="Feelings", y="Scores") +
  ggtitle("Means of Different Feelings") +
  #format of titles
  theme(legend.position="bottom",
        plot.title = element_text(
          hjust = 0.5, size = 12, face = "bold"))
```
   

   
#### Task 2: one-way ANOVA/pairwise t-tests for feelings   

##### (1) one-way ANOVA test
```{r, echo=FALSE}
###one-way ANOVA test
m1 = aov(score ~ feeling, mydata.long)
summary(m1)

###Tuckey post hoc
TukeyHSD(m1)

###Descriptive stat
plot.data
```

A one-way ANOVA shows a significant difference among the scores of feeling (F(5,288) = 53.71, *p* < .001). A post hoc analysis, together with descriptive statistics, further reveals that 1) the score of feelings_1 is significantly lower than those of feelings_2-4; 2) the score of feeling_2 is significantly higher than those of feelings_3,5,6, but significantly lower than that of feelings_4; 3) the score of feeling_2 is significantly higher than those of feelings_5,6, but significantly lower than that of feelings_4; 4) he score of feeling_4 is significantly higher than those of feelings_5,6.    

##### (2) pairwise t-tests

##### (a)feeling_1 & feeling_2
```{r, echo=FALSE}
###feeling_1 & feeling_2
feelings12 = mydata.long %>%
  filter(feeling == "feelings_1" | feeling == "feelings_2")
t.test(score ~ feeling, feelings12, paired = T, alternative = "two.sided")
```

##### (b)feeling_1 & feeling_3
```{r, echo=FALSE}
feelings13 = mydata.long %>%
  filter(feeling == "feelings_1" | feeling == "feelings_3")
t.test(score ~ feeling, feelings13, paired = T, alternative = "two.sided")
```

##### (c)feeling_1 & feeling_4
```{r, echo=FALSE}
feelings14 = mydata.long %>%
  filter(feeling == "feelings_1" | feeling == "feelings_4")
t.test(score ~ feeling, feelings14, paired = T, alternative = "two.sided")
```

##### (d) feeling_1 & feeling_5
```{r, echo=FALSE}
feelings15 = mydata.long %>%
  filter(feeling == "feelings_1" | feeling == "feelings_5")
t.test(score ~ feeling, feelings15, paired = T, alternative = "two.sided")
```

##### (e) feeling_1 & feeling_6
```{r, echo=FALSE}
feelings16 = mydata.long %>%
  filter(feeling == "feelings_1" | feeling == "feelings_6")
t.test(score ~ feeling, feelings16, paired = T, alternative = "two.sided")

```

The results of paired t-test shows that feelings_1 is significantly different from other feelings (all *p*s < .05)      



#### Task 3: create a graph showing proportion of people choosing outcome_binary1    

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE, fig.height=8, fig.width=8}
###make a proportion table
outcome_binary1 = prop.table(
  matrix(table(mydata$outcome_binary1), 
         dimnames = list(c("I apologize first, then the person apologizes", 
                           "Neither I nor the person apologizes"), 
                         "proportion")))

###change data type
outcome_binary1 = as.data.frame(outcome_binary1) #change to a data frame
outcome_binary1 = setDT(outcome_binary1, keep.rownames = T)[] #change the row name into a column
colnames(outcome_binary1)[1] = "outcome" #rename the row name
outcome_binary1$outcome = as.factor(outcome_binary1$outcome) #change to a factor
outcome_binary1$proportion = round(outcome_binary1$proportion, digits = 4)


###visualization
ggplot(outcome_binary1, aes(x="", y=proportion, fill=outcome)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  #titles
  ggtitle("Proportion of Outcome1 (%)") +
  geom_text(aes(label = proportion*100),
            position = position_stack(vjust = 0.5))+
  #format of titles
  theme(legend.position="bottom",
        plot.title = element_text(
          hjust = 0.5, size = 12, face = "bold"))
  
  
```
    



#### Task 4: proportion test of outcome_binary1   

```{r, echo=FALSE}
#10 participants choose "Neither I nor the person apologizes", while 39 choose "I apologize first, then the person apologizes"
binom.test(x = 39, n = 49, p = 0.2041)
```

A one proportion test shows that the proportion of participants who chose "I apologize first, then the person apologizes" (79.59%) is significantly higher than that of participants who chose "Neither I nor the person apologizes" (20.41%, *p* < .001).       





#### Task 5: create a graph showing proportion of people choosing outcome_binary2    

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE, fig.height=8, fig.width=8}
###similar as task 3, make a proportion table
outcome_binary2 = prop.table(
  matrix(table(mydata$outcome_binary2), 
         dimnames = list(c("I apologize first, but the person does not apologize after that",
                           "Neither I nor the person apologizes"), 
                         "proportion")))

###similar as task 3, change data type
outcome_binary2 = as.data.frame(outcome_binary2) #change to a data frame
outcome_binary2 = setDT(outcome_binary2, keep.rownames = T)[] #change the row name into a column
colnames(outcome_binary2)[1] = "outcome" #rename the row name
outcome_binary2$outcome = as.factor(outcome_binary2$outcome) #change to a factor
outcome_binary2$proportion = round(outcome_binary2$proportion, digits = 4)


###similar as task 3, visualization
ggplot(outcome_binary2, aes(x="", y=proportion, fill=outcome)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  #titles
  ggtitle("Proportion of Outcome2 (%)") +
  geom_text(aes(label = proportion*100),
            position = position_stack(vjust = 0.5))+
  #format of titles
  theme(legend.position="bottom",
        plot.title = element_text(
          hjust = 0.5, size = 12, face = "bold"))
```
    


#### Task 6: proportion test of outcome_binary2   

```{r, echo = FALSE}
#27 participants choose "Neither I nor the person apologizes", while 22 choose "I apologize first, but the person does not apologize after that"
binom.test(x = 22, n = 49, p = 0.551)
```
A one proportion test shows that the proportion of participants who chose "I apologize first, but the person does not apologize after that" (44.9%) is not significantly different from participants who chose "Neither I nor the person apologizes" (55.1%, *p* > .1)    



#### Task 7: create a single bar graph compare means of different feelings by initiator types   

```{r,echo=FALSE, fig.height=5, fig.width=12}

###change data type
mydata.long$initiator_type = as.factor(mydata.long$initiator_type)

##similar as Task 1, calculate relevant stat by feelings and initiator types
plot.data.2 = mydata.long %>%
  group_by(feeling, initiator_type) %>%
  summarise(mean = mean(score, na.rm=TRUE),
  sd = sd(score, na.rm=TRUE),
  n = n(),
  se=sd/sqrt(n),
  ci = qt(0.975, df=n-1)*se)
  
###similar as Task 1, visualize the stat but by initiator types  
ggplot(plot.data.2, aes(x=feeling, y=mean, group = factor(1))) + 
  geom_bar(stat="identity") +
  #by initiator types
  facet_wrap(vars(initiator_type))+
  #error bar
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.1) +
  #text of titles
  labs(x="Feelings", y="Scores") +
  ggtitle("Means of Different Feelings by Initiator Types")+
  #format of titles
  theme(legend.position="bottom",
        plot.title = element_text(
          hjust = 0.5, size = 12, face = "bold"))
```

